---
title: "Ilha Josina Ivermectin Trial"
subtitle: "Preliminary report: where are the roads?"
author: "Joe Brew and Carlos Chaccour"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
library(leaflet)
library(tidyverse)
library(knitr)
library(leaflet)
library(RColorBrewer)
library(sp)
library(rgeos)
library(maptools)
library(rgdal)

# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
# Basic knitr options
knitr::opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)

# # Render the how many clusters report 
# # (this is what generates the shapefiles in data products)
# # as well as the census
# this_wd <- getwd()
# setwd('../how_many_clusters/')
# render('how_many_clusters.Rmd')
# setwd(this_wd)
load('../how_many_clusters/prepared_data.RData')


# Read gpx files
# Define function for reading gpx files
read_gpx <- function(file = "../../gpx/carlos.gpx",
                     make_spatial = FALSE){
  out <- readGPS(i = "gpx", 
                      f = file, 
                      type="t")
  # Rename some columns
  out <- out %>%
    rename(date = V24,
           seconds = V17,
           longitude = V5,
           latitude = V4,
           altitude = V15,
           distance_km = V23,
           distance_mi = V22) %>%
    dplyr::select(date,
                  seconds,
                  longitude,
                  latitude,
                  altitude,
                  distance_km,
                  distance_mi) %>%
  mutate(seconds = seconds - dplyr::first(seconds)) %>%
    mutate(x = longitude,
           y = latitude,
           file_name = file)
  
  # Make spatial
  if(make_spatial){
    # coordinates(out) <- ~x+y
    x <- Line(out[,c('longitude', 'latitude')])
    x <- Lines(x, ID = 'a')
    x <- list(x)
    x <- SpatialLines(x)
    out <- x
    # out <- SpatialLinesDataFrame(sl = x,
    #                              data = data.frame(ID = 'a'),
    #                              match.ID = FALSE)
    proj4string(out) <- CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
  }
  
    
  # Return
  return(out)
}

# Get data
gpx_carlos <- read_gpx("../../gpx/carlos.gpx")
gpx_patricia <- read_gpx("../../gpx/patricia.gpx")
gpx_patricia_2 <- read_gpx("../../gpx/patricia2.gpx")
gpx_joe <- read_gpx("../../gpx/joe.gpx")
gpx_joe_2 <- read_gpx("../../gpx/joe2.gpx")

# # Lowest point
# lowest <- coordinates(gpx_carlos)[[1]][[1]][which.min(coordinates(gpx_carlos)[[1]][[1]][,2]),]

# Combine all the gpx
gpx <- bind_rows(gpx_carlos,
                 gpx_joe,
                 gpx_joe_2,
                 gpx_patricia,
                 gpx_patricia_2)
```

```{r}
library(cism)
# Get map of manhica and ilha josina
ij <- man3
ij <- ij[ij$NAME_3 == 'Ilha Josina Machel',]
ij_fortified <- broom::tidy(ij, id = 'NAME_3')

# Remove everything out of ij
# gpx_carlos = gIntersection(gpx_carlos, ij)
# gpx_patricia = gIntersection(gpx_patricia, ij)

# Make gpx spatial points
gpx_sp <- gpx
coordinates(gpx_sp) <- ~x+y
proj4string(gpx_sp) <- CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

# Remove those points outside of ij
gpx_sp <- raster::intersect(gpx_sp, ij)

# Converg back to df
gpx <- gpx_sp@data

# Clean up names
gpx <-
  gpx %>%
  mutate(Source = gsub('../../gpx/|.gpx',
                       '',
                       file_name))
gpx$Source <- Hmisc::capitalize(gpx$Source)
```

## Publicly known roads

The below map shows those roads and paths which are publicly known, via the OpenStreetMaps database. 

```{r}
library(osmar)
if('public_roads.RData' %in% dir()){
  load('public_roads.RData')
} else {
  src <- osmsource_api()
  bb <- corner_bbox(left = 32.8182,
                    bottom = -25.3782,
                    right = 33.1403,
                    top = -25.0613)
  ctown <- get_osm(bb, source = src)
  save(ctown, 
       file = 'public_roads.RData')
}

roads <- as_sp(ctown, "lines")  # convert data to sp class
# plot(ij, col = adjustcolor('red', alpha.f = 0.6), border = NA)
# lines(roads,
#       col = adjustcolor('black', alpha.f = 0.6))
# lines(gpx_carlos, col = 'red', type = 'l')
# lines(gpx_patricia, col = 'blue', type = 'l')

roads_df <- broom::tidy(roads)
# Keep only those NEAR ij
roads_df <-
  roads_df %>%
  filter(long > 32.8,
         long < 33.04,
         lat > -25.35,
         lat < -25.03)

ggplot() +
  geom_polygon(data = ij_fortified,
               aes(x = long,
                   y = lat,
                   group = group),
               color = NA, fill = 'blue',
               alpha = 0.5) +
  geom_path(data = roads_df,
             aes(x = long,
                 y = lat,
                 group = group),
            alpha = 0.6) +
  coord_map() +
  theme_cism_map()

# leaflet() %>%
#   addProviderTiles('Esri.WorldImagery') %>%
#   addPolylines(data = gpx_carlos)
```

## The problem

As can be seen above, with the exception of the northern most part, and some rivers which are incorrectly classified as roads, there exist almost no known roads.


## The solution

We travel the roads ourselves, recording our location on mobile devices. The red lines mark the areas covered so far.

```{r}

ggplot() +
  geom_polygon(data = ij_fortified,
               aes(x = long,
                   y = lat,
                   group = group),
               color = NA, fill = 'blue',
               alpha = 0.5) +
  geom_path(data = roads_df,
             aes(x = long,
                 y = lat,
                 group = group),
            alpha = 0.6) +
  coord_map() +
  geom_path(data = gpx,
            aes(x = longitude,
                y = latitude,
                group = Source),
            color = 'red') +
  theme_cism_map()
```

## Proximity to residents

The below map is identical to the above, with the one difference being that it also shows (as orange points) the location of known residents of Ilha Josina. The purpose of this map is to determine the potential need for and to guide further exploratory mapping, so as to geocode roads which are close to residents.

```{r}

ggplot() +
  geom_polygon(data = ij_fortified,
               aes(x = long,
                   y = lat,
                   group = group),
               color = NA, fill = 'blue',
               alpha = 0.5) +
  geom_path(data = roads_df,
             aes(x = long,
                 y = lat,
                 group = group),
            alpha = 0.6) +
  coord_map() +
  theme_cism_map() +
  geom_point(data = census,
             aes(x = longitude,
                 y = latitude),
             alpha = 0.3,
             size = 0.3,
             color = 'orange') +
    geom_path(data = gpx,
            aes(x = longitude,
                y = latitude,
                group = Source),
            color = 'red') 
```

## Next steps

- Further mapping of the northwest corner?  
- Calculate distance from residents to nearest known road.  

