---
title: "Ilha Josina Ivermectin Trial"
subtitle: "Preliminary report: cluster enumeration and delineation"
author: "Carlos Chaccour"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
library(leaflet)
library(tidyverse)
library(knitr)
library(leaflet)
library(RColorBrewer)
library(sp)
library(rgeos)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)
```

```{r}
# Data preparation
if('prepared_data.RData' %in% dir()){
  load('prepared_data.RData')
} else {
  source('prepare_data.R')
  save(results,
       census,
       ij,
       ij_children,
       file = 'prepared_data.RData')
}


# Get results of our simulations
x <- results %>%
  group_by(iter) %>%
  summarise(start = first(start),
            rest = first(rest),
            walk_along = first(walk_along),
            n = n(),
            ib = length(which(!is.na(cluster)))) %>%
  mutate(pib = ib / n * 100) %>%
  ungroup %>%
  mutate(p_in_core = 100 - pib) %>%
  mutate(n_in_core = n - ib)

# Get that which yields the most number of successes
best <- x %>% filter(pib == min(pib))
best <- results[results$iter == best$iter[1],]
x <- best

# x <- results
# 
# Define some colors for plotting
cols <- colorRampPalette(brewer.pal(9, 'Spectral'))(length(unique(x$cluster[!is.na(x$cluster)])))
cols <- sample(cols, length(cols))
colors <- cols[x$cluster]
colors <- ifelse(is.na(colors), 'grey', colors)

# Create a leaflet map
l <- leaflet() %>%
  # addProviderTiles('Stamen.Toner') %>%
  addTiles() %>%
  # addProviderTiles('Esri.NatGeoWorldMap') %>%
  addCircleMarkers(lng = x$lon,
                   lat = x$lat,
                   color = colors,
                   radius = 0.2,
                   popup = paste0('Cluster ', x$cluster))

# Add polygons for convex hull and for buffer zone
buffer <- 1000
for (i in 1:length(unique(x$cluster[!is.na(x$cluster)]))){

  this_cluster <- x %>% filter(cluster == i)
  this_color <- cols[i]
  lf <- this_cluster
  coordinates(lf) <- ~lon+lat
  proj4string(lf) <- proj4string(ij)
  lf <- rgeos::gConvexHull(lf)
  # Transform to projected form
  lfp <- spTransform(lf, CRS( "+init=epsg:3347" ) )
  # Get buffer at 1km
  lfb <- rgeos::gBuffer(lfp, width = buffer)
  # Un -project lfb
  lfb <- spTransform(lfb, proj4string(ij))
  lfp <- spTransform(lfp, proj4string(ij) )
  
  if(!class(lfp)[[1]] %in% c('SpatialPoints', 'NULL')){
      l <- l %>%
    addPolylines(data = lfp,
                 color = this_color,
                 weight = 2,
                 fill = TRUE) %>%
    addPolylines(data = lfb,
                 color = this_color,
                 weight = 1)
  }
}

```

# Context

This document is a first step in analyzing the feasibility of carrying out an Ivermectin trial in the area of Ilha Josina, District of Manhiça, Province of Maputo, Mozambique. Its objective is to quantify how many clusters of 50 children (age < 5) could be formed, if each of those clusters had a "buffer" of 1 kilometers (ie, minimum distance of 2 kilometers) between two children of two different buffers.

# Methods

Census data were used, with a snapshot from May 6, 2016. Residents of Ilha Josina were identified through polgyonal overlay with publicly avaiable administrative shapefiles. Those who were < 5 years of age as of the snapshot date were considered eligible. 

A custom algorithm was devised for division into clusters. The algorithm took as parameters multiple different "strategies" for creating clusters (centroid-based, "random walk" based, etc.). A simulations-based approach was used to identify the optimal strategy for the case of Ilha Josina.^["Optimal" being the maximum number of children/clusters meeting the buffering and age requirements.]

Full code for the algorithm and its implementation is available online.^[https://github.com/joebrew/ilha_josina]

# Results  

Of the `r nrow(ij_children)` residing in Ilha Josina as of the time of our snapshot, the maximum number of 50-child clusters we can form is `r max(best$cluster, na.rm = TRUE)` (ie, `r length(which(!is.na(best$cluster) & !best$in_buffer))` children). 

Our cluster centroids can be viewed below.

```{r}
centroids <- best %>%
  filter(!is.na(cluster)) %>%
  group_by(cluster) %>%
  summarise(lon = mean(lon),
            lat = mean(lat))
plot(ij)
points(centroids$lon,
       centroids$lat,
       col = cols)
```

The below plot shows the actual boundaries of our clusters, with buffers.

```{r}
plot(ij)

# Add polygons for convex hull and for buffer zone
buffer <- 1000
for (i in 1:length(unique(x$cluster[!is.na(x$cluster)]))){

  this_cluster <- x %>% filter(cluster == i)
  this_color <- cols[i]
  lf <- this_cluster
  coordinates(lf) <- ~lon+lat
  proj4string(lf) <- proj4string(ij)
  lf <- rgeos::gConvexHull(lf)
  # Transform to projected form
  lfp <- spTransform(lf, CRS( "+init=epsg:3347" ) )
  # Get buffer at 1km
  lfb <- rgeos::gBuffer(lfp, width = buffer)
  # Un -project lfb
  lfb <- spTransform(lfb, proj4string(ij))
  lfp <- spTransform(lfp, proj4string(ij) )
  
  if(!class(lfp)[[1]] %in% c('SpatialPoints', 'NULL')){
    plot(lfp, add = TRUE, col = cols[i])
    plot(lfb, add = TRUE,
         col = adjustcolor(cols[i], alpha. = 0.6),
         border = NA)
  }
}

```

The below interactive map shows all clusters, buffers, and child residents. Points are children, colored by their respective clusters. Grey points are children residing outside of core cluster areas (ie, buffers or unclusterable). Thick lines show cluster borders; thin lines show cluster buffer boundaries.

```{r}
l
```

# Assignation to intervention


```{r}
# Create a dataframe of clusters
clusters <- results %>%
  group_by(cluster) %>%
  summarise(x_centroid = mean(x),
            y_centroid = mean(y)) %>%
  filter(!is.na(cluster))

# Set a random seed
set.seed(510)

# Create assignations
clusters$status <- sample(c('intervention',
                                'control'),
                              nrow(clusters), 
                              replace = TRUE)
```

We randomly assign all `r nrow(clusters)` clusters to be of either "intervention" or "control" status. The assignments can be seen below.

```{r}
kable(clusters %>%
        dplyr::select(-x_centroid,
                      -y_centroid))
```

# Voronoi tesselation

In order both to (a) expand our cluster boundaries to the maximum possible size and (b) create "implementation" zones (ie, areas defined by whether they receive the intervention or not), we perform Delaunay triangulation / Voroni tesselation.


```{r}
# Make a spatial version of results
results_sp <- results %>% filter(!is.na(cluster))
coordinates(results_sp) <- ~x+y

# Create convex hulls
ch <- list()
for (i in 1:nrow(clusters)){
  this_cluster <- clusters$cluster[i]
  sub_results_sp <- results_sp[results_sp@data$cluster == this_cluster,]
  x <- rgeos::gConvexHull(sub_results_sp)
  ch[[i]] <- x
}

# Create delaunay triangulation / voronoi tiles for entire surface
voronoi <- function(shp = results_sp){
  
  # shp@data <- data.frame(shp@data)
  # 
  # # Fix row names
  # row.names(shp) <- 1:nrow(shp)
  
  # Remove any identical ones
  shp <- shp[!duplicated(shp$lon, shp$lat),]
  
  # Helper function to create coronoi polygons (tesselation, not delaunay triangles)
  # http://carsonfarmer.com/2009/09/voronoi-polygons-with-r/
  voronoipolygons = function(layer) {
    require(deldir)
    crds = layer@coords
    z = deldir(crds[,1], crds[,2])
    w = tile.list(z)
    polys = vector(mode='list', length=length(w))
    require(sp)
    for (i in seq(along=polys)) {
      pcrds = cbind(w[[i]]$x, w[[i]]$y)
      pcrds = rbind(pcrds, pcrds[1,])
      polys[[i]] = Polygons(list(Polygon(pcrds)), ID=as.character(i))
    }
    SP = SpatialPolygons(polys)
    voronoi = SpatialPolygonsDataFrame(SP, data=data.frame(x=crds[,1], 
                                                           y=crds[,2], row.names=sapply(slot(SP, 'polygons'), 
                                                                                        function(x) slot(x, 'ID'))))
  }
  # http://gis.stackexchange.com/questions/180682/merge-a-list-of-spatial-polygon-objects-in-r
  appendSpatialPolygons <- function(x) {
    ## loop over list of polygons
    for (i in 2:length(x)) {
      # create initial output polygon
      if (i == 2) {
        out <- maptools::spRbind(x[[i-1]], x[[i]])
        # append all following polygons to output polygon  
      } else {
        out <- maptools::spRbind(out, x[[i]])
      }
    }
    return(out)
  }
  
  tile_polys <- voronoipolygons(shp)
  # Add the cluster numbers
  tile_polys@data$cluster <- the_clusters <- shp$cluster
  cols <- rainbow(as.numeric(factor(tile_polys@data$cluster)))
  
  # Disolve borders
  x = gUnaryUnion(tile_polys, id = tile_polys$cluster)
  
  jdata = SpatialPolygonsDataFrame(Sr=x, 
                                   data=data.frame(cluster = as.numeric(as.character(names(x)))),FALSE)
  
  return(jdata)
}

# Get voronoi tesselations
results_spv <- voronoi(shp = results_sp)

```


Having carried out the polygon delineation, we plot the clusters with expanded borders.

```{r}
# Plot in current state
plot(results_spv)
plot(ij, add = T, col = adjustcolor('blue', alpha.f = 0.2))
```

In the above, our methodology works well within Ilha Josina; but at borders, our neighborhoods expand infinitely outwards. In order to account for this, we can trim our polygons to keep only those segments which fall into Ilha Josina.

```{r}
# Narrow down so as to only keep those areas which are IN Magude
proj4string(results_spv) <- proj4string(ij)
out <- gIntersection(results_spv, ij, byid=TRUE)

# Join with data
row.names(out) <- as.character(1:length(out))
out <- SpatialPolygonsDataFrame(out, data.frame(clusters), match.ID = FALSE)

# Plot again
plot(ij, col = adjustcolor('red', alpha.f = 0.2))
plot(out, add = T, lwd = 0.2)
```

The below interactive map allows for manual examination of clusters. Click on any polygon to reveal its cluster number and intervention status.

```{r}

out@data$color <- ifelse(out@data$status == "control", "green",
                         ifelse(out@data$status == "intervention", "red", "grey"))
# Plot leaflet
library(leaflet)
leaflet() %>%
  addProviderTiles(provider = 'Stamen.Toner') %>%
  addPolygons(data = out,
              color = out@data$color,
              popup = paste0('Cluster ', as.character(out@data$cluster),
                             '. Status: ', as.character(out@data$status))) 

```

# Defining intervention zones.

Having expanded borders, we can re-calculate buffers based not on cluster number but on intervention status (since there is no need for a buffer between two clusters of identical intervention status).

To do this, we first "dissolve" our polygons by intervention status. This means removing boundaries between areas of the same intervention status. This results in the following intrevention zones (red = intervention; green = control).

```{r}
collapsed <- gUnaryUnion(out, id = out@data$status)

collapsed <- SpatialPolygonsDataFrame(collapsed, data.frame(zone = 1:length(collapsed@polygons)), 
                                      match.ID = FALSE)

# Get the intervention status
tester <- clusters %>% filter(!duplicated(status))
coordinates(tester) <- ~x_centroid+y_centroid
proj4string(tester) <- proj4string(collapsed)
x <- over(tester, polygons(collapsed))
tester$zone <- collapsed@data$zone[x]
tester <- tester@data %>%
  dplyr::select(zone, status)

# Join the intervention status to collapsed
collapsed@data <-
  left_join(x = collapsed@data,
            y = tester,
            by = 'zone')

# Add a color
collapsed@data$color <- ifelse(collapsed@data$status == 'control',
                               'green',
                               ifelse(collapsed@data$status == 'intervention',
                                      'red', 'grey'))

leaflet() %>%
  addProviderTiles(provider = 'Stamen.Toner') %>%
  addPolygons(data = collapsed,
              color = collapsed@data$color)

```

We then use spatial buffering to re-define our buffers, going inwards 1,000 meters from the "dissolved" boundaries.

```{r}
# Transform to projected form
collapsed_projected <- spTransform(collapsed, CRS( "+init=epsg:3347" ) )
# Get buffer at 1km
collapsed_projected_buffered <- rgeos::gBuffer(collapsed_projected, 
                                               width = -1000,
                                               byid = TRUE)
# Un -project collapsed_projected_buffered
collapsed_buffered <- spTransform(collapsed_projected_buffered, proj4string(ij))
leaflet() %>%
  addProviderTiles(provider = 'Stamen.Toner') %>%
  addPolygons(data = collapsed_buffered,
              color = collapsed_buffered@data$color)

```


## Data products

Having finished our implementation of cluster delineation, we write software-agnostic shapefiles. The purpose of this step is to remove the data products (cluster and buffer delineations) from this report, so that others can use these clusters, regardless of their GIS software.

We create 3 data products.

### 1. `ij_clusters`

`ij_clusters` is a projected shapefile with the exact boundaries between clusters. It is downloadable from: https://github.com/joebrew/ilha_josina/tree/master/data_products/ij_clusters

```{r}
plot(out,
     col = rainbow(nrow(out)))
library(rgdal)
owd <- getwd()
setwd('../../data_products/')
writeOGR(obj=out, 
         dsn="ij_clusters", 
         layer = 'ij_clusters',
         driver="ESRI Shapefile",
         overwrite_layer = TRUE)
setwd(owd)
```

### 2. `ij_intervention_zones`

`ij_intervention_zones` is a projected shapefile with the exact boundaries of the intervention zones. It is downloadble from: https://github.com/joebrew/ilha_josina/tree/master/data_products/ij_intervention_zones

```{r}
plot(collapsed,
     col = collapsed@data$color)
owd <- getwd()
setwd('../../data_products/')
writeOGR(obj=collapsed, 
         dsn="ij_intervention_zones", 
         layer = 'ij_intervention_zones',
         driver="ESRI Shapefile",
         overwrite_layer = TRUE)
setwd(owd)
```


### 3. `ij_intervention_zones_buffered`

`ij_intervention_zones_buffered` is a projected shapefile with the exact boundaries of the intervention zones with a -1000 meter inward buffer added to each zone. It is downloadable from: https://github.com/joebrew/ilha_josina/tree/master/data_products/ij_intervention_zones_buffered

```{r}
plot(collapsed_buffered,
     col = collapsed_buffered@data$color)
owd <- getwd()
setwd('../../data_products/')
writeOGR(obj=collapsed_buffered, 
         dsn="ij_intervention_zones_buffered", 
         layer = 'ij_intervention_zones_buffered',
         driver="ESRI Shapefile",
         overwrite_layer = TRUE)
setwd(owd)
```

# Increased catchment


```{r}
# Get zones
results_sp <- results
coordinates(results_sp) <- ~x+y
proj4string(results_sp) <- proj4string(ij)
x <- over(results_sp, collapsed_buffered)
results_sp@data <- cbind(results_sp@data, x)

# See which ones now get a zone
results_sp@data <-
  results_sp@data %>%
  mutate(new_zone = is.na(cluster) & !is.na(zone))
```




Though our dissolution method does not increase our number of clusters, it _does_ increase the catchment of each area. Whereas our previous population size was `r nrow(clusters) * 50`, having dissolved unecessary boundaries, we increase our estimated non-buffer population size by `r length(which(results_sp@data$new_zone))` to `r length(which(results_sp@data$new_zone)) + (nrow(clusters) * 50)`. In reality, given population growth since the time of the census, this figure may even be higher.

The below map shows intervention assignments. Small dots represent children. Small dots in a triangle represent children who were originally considered to be in buffer areas (when buffers were drawn at the cluster level), but are now in "core" areas (when buffers are drawn at the intervention zone level).

```{r}
results_sp@data$color <- 
  ifelse(is.na(results_sp@data$color),
         'blue',
         results_sp@data$color)
plot(ij)
points(results_sp, 
     col = results_sp@data$color,
     pch = '.')

new_zone <- results_sp[which(results_sp@data$new_zone),]
points(new_zone, 
     col = new_zone@data$color,
     pch  = 2,
     cex = 0.5)
```


# Details

All code for the methods discussed in this report is downloadable at: https://github.com/joebrew/ilha_josina/tree/master/reports/how_many_clusters